A short introduction to causal inference

### 2015-10-22 causality_tutorial.tex
lots of changes - laptop has battery issues
prep for tpr talk

TODO
1BVAEU

@@

The queries we have been asking thus far are about $P(Y^{0}=1) = \gamma + \delta$ and $P(Y^{1}=1) = \beta + \delta$, but suppose we asked the question; what is the probability that this patient, who was not treated and died, would have recovered if they had been treated? We know they are in either group 1 or 2 since they died without treatment, so the answer is $\frac{\beta}{\alpha+\beta}$. Can we estimate the $\alpha, \beta, \gamma, \delta$ or in other words, identify the joint distribution over the counterfactuals $P(Y^{0},Y^{1})$ given the interventional distributions, $P(Y^{0})$ and $P(Y^{1})$? The answer is no, putting our constraints and unknowns in matrix form:D-separation still applies in the augmented model.

@@

>>>>

@@

The queries we have been asking thus far are about $P(Y^{0}=1) = \gamma + \delta$ and $P(Y^{1}=1) = \beta + \delta$, but suppose we asked the question; what is the probability that this patient, who was not treated and died, would have recovered if they had been treated? We know they are in either group 1 or 2 since they died without treatment, so the answer is $\frac{\beta}{\alpha+\beta}$. Can we estimate the $\alpha, \beta, \gamma, \delta$ or in other words, identify the joint distribution over the counterfactuals $P(Y^{0},Y^{1})$ given the interventional distributions, $P(Y^{0})$ and $P(Y^{1})$? The answer is no, putting our constraints and unknowns in matrix form:

@@

A comparison of algorithms

\renewcommand{\arraystretch}{1.5}
\begin{tabular}{| c | c | p{4cm} | c | c | c |}
\hline
  \textbf {Alg.} &\textbf{ Method }& \textbf{Scales (num.vars) }& $\sim $ \textbf {Vars} & \textbf {Latent } & \textbf {Reference} \\
  \hline
  IC/SGS & Constraint based & Exponential & 10 & No & Pearl(2000)/Sprites(2000)\\
  \hline
  PC & Constraint based & Worst case exponential, polynomial for sparse graphs & 5000 & No & Sprites(2000) \\ 
  \hline
  FCI & Constraint based & Worst case exponential, polynomial variant FCI+ for sparse graphs & 30 & Yes & Sprites(2000) \\
  \hline
  RFCI & Constraint based & ? & 500 & Yes & Colombo(2012) \\
  \hline
  GES & Search \& Score  & Worst case exponential & 50 & No & Chickering(2002) \\
  \hline
  MMHC & Hybrid & ? & 5000 & No & Tsamardinos(2006) \\
 \hline
\end{tabular}

### 2015-10-22 graph_diagram_cholesterol.tex
prep for tpr talk

\documentclass{article}
\usepackage{tikz} 
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{algorithm}% http://ctan.org/pkg/algorithm
\usepackage[noend]{algpseudocode}% http://ctan.org/pkg/algorithmicx

\usepackage{tikz}

@@

\tikzset{
    %Define standard arrow tip
    >=stealth',
    %Define style for boxes
    observed/.style={
           circle,
           rounded corners,
           draw=black, thick,
           minimum width=.5em,
           minimum height=.5em,
           font=\footnotesize,
           text centered,
           fill=blue!20!white},
     latent/.style={
           circle,
           rounded corners,
           draw=black, thick, dashed,
           minimum width=.5em,
           minimum height=.5em,
           font=\footnotesize,
           text centered,
           fill=black!10!white
           },
    % Define arrow style
    pil/.style={
           o->,
           thick,
           shorten <=2pt,
           shorten >=2pt,},
    sh/.style={ shade, shading=axis, left color=red, right color=green,
    shading angle=45 }

@@

\begin{document}
\def\ci{\perp\!\!\!\perp} % from Wikipedia

\begin{figure}
\centering
\begin{subfigure}[t]{0.44\textwidth}
\centering
\caption{}
\begin{tikzpicture}[->,shorten >=0pt,shorten <=0pt,node distance=0.45cm,  thick,main node/.style={observed}, shaded/.style={latent}]
\node[main node](1){C};
\node[main node, above left=of 1](2){A};
\node[main node, above right=of 1](3){B};
\node[main node, below=of 1](4){D};

 \path[]
 	(1) edge node {} (4)
    (2) edge [->]  (1)
    (3) edge [->]  (1);

\node[main node,right of=3,xshift=1.2em](5){A};
\node[main node, below right=of 5](6){C};
\node[main node, above right=of 6](7){B};
\node[main node, below=of 6](8){D};
\path[every node/.style={font=\sffamily\footnotesize}]
	(1) edge [-, >=stealth', blue!30!white, line width=.9em,shorten <=1em,shorten >= 1.7em,
	preaction={line width = .3em, blue!30!white,-triangle 90,draw,shorten >=.6em,shorten <=1em}
	] node[black]{FCI} (6)
	(5) edge [o->] (6)
	(7) edge [o->] (6)
	(6) edge [->] (8);

\end{tikzpicture}
\end{subfigure}
\begin{subfigure}[t]{0.55\textwidth}
\centering
\caption{}
\begin{tikzpicture}[->,shorten >=0pt,shorten <=0pt,node distance=0.45cm,  thick,main node/.style={observed}, shaded/.style={latent}]

\node[main node](1){A};
\node[main node, below right=of 1](2){B};
\node[shaded, above right=of 2](3){U};
\node[main node, below right=of 3](4){C};
\node[main node, above right=of 4](5){D};
\node[ below=of 3,yshift=.2cm](11){};
\path[] (1) edge (2)
		(3) edge (2) edge (4)
		(5) edge (4);

\node[main node,right=of 5,xshift=-.5em](6){A};
\node[main node, below right=of 6](7){B};
\node[above right=of 7](8){};
\node[main node, below right=of 8](9){C};
\node[main node, above right=of 9](10){D};
\node[below = of 8,yshift=0.2cm](12){};	
\path[every node/.style={font=\sffamily\footnotesize}]
(11) edge [-, >=stealth', blue!30!white, line width=.9em,shorten <=3.3em,shorten >= 4.5em,
	preaction={line width = .3em, blue!30!white,-triangle 90,draw,shorten >=3.7em,shorten <=4em}
	] node[black,xshift=-.5em]{FCI} (12)
(6) edge [o->] (7)
(7) edge [<->] (9)
(10) edge [o->](9);

\end{tikzpicture}
\end{subfigure}
\end{figure}

>>>>

\documentclass{article}
\usepackage{tikz} 
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{algorithm}% http://ctan.org/pkg/algorithm
\usepackage[noend]{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage[margin=0.7in]{geometry}
\usepackage{tikz}

@@

\tikzset{
    %Define standard arrow tip
    >=stealth',
    %Define style for boxes
    observed/.style={
           rectangle,
           rounded corners,
           draw=black, thick,
           minimum width=6em,
           minimum height=2.2em,
           font=\footnotesize,
           text centered,
           },
     latent/.style={
           circle,
           rounded corners,
           draw=black, thick, dashed,
           minimum width=2.2em,
           minimum height=2.2em,
           font=\footnotesize,
           text centered,
           fill=black!10!white
           },
    % Define arrow style
    pil/.style={
           o->,
           thick,
           shorten <=2pt,
           shorten >=2pt,},
    sh/.style={ shade, shading=axis, left color=red, right color=green,
    shading angle=45 }

@@

\begin{document}
%
\def\ci{\perp\!\!\!\perp} % from Wikipedia

\begin{figure}
\centering
\begin{tikzpicture}[->,shorten >=0pt,shorten <=0pt,node distance=3em,thick,main node/.style={observed}, lt/.style={latent}]
\node[main node](1){Health concious};
\node[main node, below left=of 1](2){Diet};
\node[main node, below right=of 1](3){Exercise};
\node[main node, below right=of 2](4){Weight};
\node[main node, fill=black!20!white,below left=of 4](5){Y=Cholesterol};
\node[main node, below =of 5](6){Test result};
\node[main node, above right=of 3](7){Free time};
\path[]
	(1) edge (2) edge (3)
	(2) edge (4) edge (5)
	(3) edge (4)
	(5) edge (6)
	(7) edge (3);
\end{tikzpicture}
\end{figure}
